How to Learn Useful Changes to Software Projects\\(to Reduce Runtimes and Software Defects)

### 2015-08-28 Manuscript.tex
Updates from ShareLaTeX

 This paper proposes  the  XTREE planner for software projects. 
XTREE inputs tables of data of independent features plus a   weighted class
  indicating how much each row is  ``bad'' or  ``better''. 
  Plans are edits to  
rows such that the changed row is more likely to be ``better''.  XTREE
learns those plans by building a decision tree across the data, then reporting
the differences in the branches from some current branch to another desired branch.

Using data from 11 software projects,  XTREE found   larger reductions than three
alternate methods.  Those reductions had 
median size of (56\%,28\%) and   largest size of (60\%,77\%) in
(defects,runtimes), respectively.
\end{abstract}
\section{Introduction}
Business users are demanding  tools that support    business-level interpretations of their data. At a panel on software analytics at ICSE`12, industrial practitioners lamented the state of the art in software analytics~\cite{menzies12a}. Panelists commented  ``prediction is all well and good, but what about decision making?''. Note that these panelists were more interested in the interpretations and follow-up that occurs after the mining, rather than just the mining itself. So:
\bi
\item
Instead of just accepting {\em predictions} on how many software defects to expect, business users might now demand a {\em plan} to reduce the likelihood of those defects.
\item
Instead of just accepting {\em predictions} on the runtime time of their software, business users might now demand a {\em plan} to reduce that runtime. 
\ei
In response to this business-level demands for planners,
we propose a novel {\em planning} method called XTREE for learning changes to a software system
such that its performance ``improves'', according to some measure. This paper uses XTREE
to reduce  the expected value of the
defects     in Jureczko et al.'s    JAVA systems~\cite{jureczko10};
and the runtimes   in   software    configured by  Siegmund et al.~\cite{sven12}.

@@

The rest of this paper   describes our data, our planners, and the experiment that
ranks XTREE against alternate approaches.  This is followed by notes on related work
and validity.
To enable reproducibility, all scripts and data used in this 
study are available on-line at github.com/ai-se/XTREE.

@@

\subsection{From Prediction to Planning}

@@

One way to propose changes to a project would be to   ask some smart experienced
person for their opinion on how to (e.g.) reduce defects and/or decrease runtimes. Sometimes
such advice 
is an effective strategy and sometimes it is not.
According to Passos et al.~\cite{passos11},  developers
may  assume that the lessons they learn from a few past
projects are general to 
all their future projects. They comment ``past experiences were taken into account without 
much consideration for their context''~\cite{passos11}.  
 J{\o}rgensen \& Gruschke~\cite{jorgensen09} offer  a similar warning. They report that 
  supposed software engineering    ``Gurus'' rarely use lessons
  from past projects to improve their future reasoning and that such poor
  past advice can be detrimental to new projects.~\cite{jorgensen09}.
  Accordingly, we   propose a ``trust, but verify'' approach.
  After a software guru offers some sage wisdom,  it is wise to ask some other oracle 
  if there are any better options
  (just as a sanity check).
  The rest of this paper discusses some methods to build automatic oracles 
  to implement that   sanity check.

Another way  to find   changes to a project
might be to rely
on the peer review processes used by the 
SE research community. This approach would propose changes to software
projects that concur with internationally accept best practices. 
There are two
problems with that approach. Firstly, given the rapid pace of change in software
engineering, we may be asking questions for which there is no currently acceptable
``best practice''. For example, only very recently has there been any work
of predicting software runtimes based on choices within Makefiles. While the work
of Siegmund et al.~\cite{sven12} is certainly state of the art, it does not yet represent ``internationally accept best practices'' since that work is so
recent.

@@

\subsection{Trusting the Changes}\label{sect:trust}
   XTREE is evaluated by  comparing
predicted performance scores before and after a planner makes changes to the feature values of an example:
After making those
changes, we may have a new example that has never seen before. Therefore, it must be asked
{\em ``can we trust the predictions made on such new examples?''}

@@

Accordingly, we should use {\em trust-increasing} planners that generate changed examples {\em closer} to the
training examples.  To see how this works, 
 \fig{howxy} is from the {\em ivy} data
set, which is one of the Jureczko data sets explored in this paper. It shows: (1)~the training examples in gray, (2)~the test examples in red, and (3)~the
changed  examples displaced after applying a plan (in green).
 Note that the  the   changed examples
cases  (shown in green)  fall closer to the training cases (shown in gray) than
the test cases (shown in red).

@@

 \section{Test Data}\label{sect:tesd}
To assess our planning methods, we use data from Jureczko et al.'s object-oriented JAVA systems~\cite{jureczko10}  and
  software system   configuration data from by  Siegmund et al.~\cite{sven12}\footnote{ See \url{http://git.io/vGYxc} for full access to this data.}.

   The Jureczko data records number of known defects for each class using a post-release bug tracking system. The classes are described in terms of
  nearly two dozen metrics such as number of children (noc), lines of code (loc), etc. For details on the Jureczko data, see  \fig{ck} and \fig{jd}. 
 All the planning methods of this paper reflect on the numeric value of the raw defect counts. The predictor considers defects to a Boolean class data set, where defects are \texttt{TRUE} if the numeric defect count is greater than zero and \texttt{FALSE} otherwise.

@@

  Our evaluation strategy (discussed below) divides this data into a training and a test set.
  From the train set we apply a data miner (to learn a quality predictor) and
  various planning methods (to learn different plans). Next, we try applying
  each of those plans to the  test set and ask the quality predictor to assess the changed examples.
  Finally, we say that the  ``best'' planner is the one that most reduces the predicted values
  in the changed examples.

  As mentioned in the last section,  this approach depends on having effective predictors for assessing the results.
  For the Siegmund data, this criteria was   relatively easy to achieve.
  The data in those data sets have a continuous class (runtime of the compiled system)
  so the performance of a quality predictor can  be measured in terms of  difference between the predicted runtime $p$ of test case items and their actual runtimes $a$ using  $s= (1 - \frac{abs(a - p)}{a})\times100\%$ ({\em higher} values are {\em better}).
This paper  explores six Siegmund configuration data sets:  Berkeley DB (Java and C versions), Apache, SQLite, LLVM, and
  x264. 
  As a preliminary study, we split that data into equal sized train:test groups
  and trained a Random Forest
  Regressor (from the SciKit learn kit~\cite{Pedregosa2012})   on one half, then applied to the other. After repeating this process 40 times, we achieved nearly perfect accuracy of $s=\{99.9, 99.8, 99.4, 99.1, 96.1\}\%$.
Therefore, we can be very confident that the predictors from the Siegmund data can assess
our plans. (Aside: if the reader doubts that such high scores are achievable, we note that these scores are consistent with those achieved by predictors built by Siegmund et al.~\cite{sven12}.)

@@

As a preliminary study, we split the Jureczko  data into train and test groups.
Random Forests (again, from the SciKit learn kit~\cite{Pedregosa2012}) were
built from the training data, then applied to the test data.
The ``untuned'' columns of \fig{j} shows those results. We called a data set ``usable'' if Random Forest was able to classify majority of the instances correctly. For this purpose, we set a threshold of $\mathit{pd}\ge 60 \wedge \mathit{pf} \le 30$\% to select suitable data sets. With this threshold, however, none of our data sets were be suitable for this study.

@@

In summary, while we cannot trust predictors from some of our defect data sets,
we can plan ways to reduce defects in {\em jedit, ivy, ant, lucene} and {\em poi}.
Accordingly, when this study explores the Jureczko data, we will use these five data sets.

@@

This section described XTREE (which we call Method4) and  three  
alternate methods for learning plans.
  XTREE  uses the decision tree learner of \fig{where}.D.  It is new to this paper.
  The other methods use the   top-down
	bi-clustering method described in \fig{where}.C  which recursively divides the
	data in two  using a dimension that captures the greatest variability in the data. 
	We proposed Method 1 and 2   in 2012~\cite{me12c} while Methods 3 comes from research conducted earlier this year~\cite{krishna15}.

  Note that all   methods have the  properties proposed in \tion{prelim}.
  That is, they are {\em local learners}; i.e. different test examples
  will be given plans that are specialized  to their particulars. Also,
  they  are {\em trust-increasing}; 
  i.e. they change  examples such that they move {\em closer} to the training data.

@@

Our  description of the methods adopts the following convention. All variables
  set via  our engineering judgement  with Greek letters; e.g. $\alpha,\beta,\gamma$.
  In this paper, we show our current settings to these variables produces useful
  results. Elsewhere\cite{krall14,fu:ase15}, we are exploring tuning methods to 
  find better settings but  we have nothing definitive yet to report
  on auto-tuning planners.

@@

{\em Assumption1: } Method1 assumes that large data sets can be adequately represented by a few dozen (or so) 
centroids.

 {\em Details1:} Method1 clusters project data by reflecting   on the independent variables, then
  reports the delta between the cluster centroids. 
  After clustering   training data using the WHERE algorithm of  \fig{where}.C, Method1
replaces all clusters with a  centroid $C_i$ computed from the mean/mode value of each
continuous/discrete feature. After that, it
finds the closest centroid $C_j$ that has a better
performance score. For defect data, ``better'' means fewer defective examples while for the config data,
``better'' means lower median runtimes for the examples in that cluster.
Method1 then caches the  delta between the independent features between $C_i$ and $C_j$. For continuous
features, this delta is $C_j - C_i$. For discrete values, this delta is the value of that feature
in $C_j$. 
Finally, for every test case, Method1 use the distance measure $d$ shown in \fig{where}.B to find
the nearest centroid $C_i$.  It then proposes a plan for improving that test case
that is the conjunction of all the deltas between $C_i$ and $C_j$.

@@

{\em Summary2:} Method2 works line Method1 but now the   plans  only
mention the $\beta=33\%$ most informative features. Hence, Method2's plans are simpler.

{\em Assumption2: } Method2 assumes that, when reasoning about centroids, we can
just use   features
that   best distinguish    centroids; i.e. whose values appear in just a few centroids.

@@

 \input{tex/algos}
\subsubsection{Method3= GD=   Method2 + Gradient}\label{sect:BIC}

{\em Summary3:} Method3 is like Method2, but it uses more knowledge about the training data.

{\em Assumption3: } Method3 assumes that there exists ``gradients'' between and
within clusters which, if used, will better guide us to finding better plans.

{\em Details3:}
Method3 summarizes clusters into {\em two} examples: (1)~the centroid $C_x$ found in Method1 and
(2)~the best-in-cluster $B_x$  example; i.e. the  example in that cluster
with the best performance score. 
For the Siegmud data, $B_x$ is the cluster member with the fastest runtime;
for the Jureczko data, $B_x$  is the cluster member with lowest raw defect count
(resolving ties at random).

Method3 connects  each centroid to a nearest neighbor
by a {\em gradient}.
Each gradient has a (bottom,top) end labelled  ($C_i,C_j$) containing the  (worst,best) centroid performance scores, respectively..  
For each test instance, Method3 
find the nearest gradient, the runs up to the top  best end $C_j$, then extracts $B_j$ (which is the
 best-in-cluster associated with  $C_j$).
The returned plan is then computed as a fraction of the gradient between the test case
and $B_j$.

@@

 \begin{figure}[!h]
{\small 
\[
\begin{array}{r} 
\mathrm{project}\\
\mathrm{data}
\end{array} 
\left\{\begin{array}{l}\mathit{train}
        \left\{\begin{array}{l}
                \mathrm{learn\;a\;}\mathrm{predictor\;}\mathrm{(e.g.\;via \;Random\;Forest)}\\
                \mathrm{learn\;a\;}\mathrm{planner\;}\mathrm{(e.g.\;via \; XTREE)}
              \end{array}\right.
       \\
      ~\\
\mathit{test}  
    \left\{\begin{array}{l@{~}l}
           \mathit{before}& =\mathrm{Performance\; scores \hspace{2pt} in \hspace{2pt}}\mathit{test}\\
           \mathrm{\bf if\;}\mathit{before} & >  \mathit{0}\\
           \mathrm{\bf then} &
           \left\{
            \begin{array}{l}
                \mathit{test'} = \mathrm{planner}(\mathit{test})\\
                \mathit{after} =\mathrm{predictor}(\mathit{test'})\\ 
                \mathrm{{\bf return}\;} R=(1-\frac{\mathit{after}}{\mathit{before}})\times 100 \%
            \end{array}
          \right.
   \end{array}\right.
\end{array} \right. 
\]}
 \caption{Experimental design .}\label{fig:design}
 \end{figure}

@@

Else, we (1)~apply the { planner} to alter the {\em test} data;
then (2)~apply the { predictor} to the altered data $test'$;
then (3)~return data on the {\em before, after} predictions expressed as percent improvement $R=(1-\frac{\mathit{after}}{\mathit{before}})\times100\%$ with the following following properties:
\bi
\item If $R= 0\%$, this means  ``no change from baseline''; 
\item If $R < 0\%$, this indicates ``some reduction to the baseline'';
\item If $R > 0\%$, this indicates ``optimization failure''.
\ei
 \subsubsection{Statistical Methods}
 Our methods use some stochastic algorithms; e.g. WHERE's selection of ``what example to explore first'' (see \fig{where}.C) and
  XTREE' occasional use of a random guess when deciding what part of a discretized range to include in the plan
  (see \fig{xtrees_bare}). Hence, we report the $R$ values seen in 40 repeated runs
  (with different random number seeds).
The the value 40 was chosen to be  larger than the 30 samples  required
to satisfy the central limit theorem.

  To rank our methods using the results from these 40
  repeats, we use the Scott-Knott test recommended by Mittas and  Angelis~\cite{mittas13}. 
In that test, using the median values of each method,  
sort a list of  $l=40$ values of $R$ values found in  $ls=4$ different methods. 
Then,
splits $l$ into sub-lists $m,n$ in order to maximize the expected value of
 differences  in the observed performances
before and after divisions. E.g. for lists $l,m,n$ of size $ls,ms,ns$ where $l=m\cup n$:
 \[E(\Delta)=\frac{ms}{ls}abs(m.\mu - l.\mu)^2 + \frac{ns}{ls}abs(n.\mu - l.\mu)^2\]
Scott-Knott  applies an statistical  hypothesis test $H$ to check
if $m,n$ are significantly different  (in our case, the conjunction of A12 and bootstrapping). 
If so, Scott-Knott  recurses on the splits.
In other words, the Scott-Knott procedure being used here divides the data if \textit{both} bootstrap sampling and effect size test agree that a division is statistically significant (with a confidence of 99\%) and not a small effect ($A12 \ge 0.6$).

@@

Our results are presented by the line diagrams like those shown on the right-hand-side of the following example table.
The black dot shows the median $R$ value and the horizontal likes stretch over the inter-quartile
range (hereafter, IQR) that is the space from the 25th percentile value to
the 75th percentile value.

@@

{\small  \begin{tabular}{{l@{~~~~}l@{~~~~}r@{~~~~}r@{~~}c@{}r}} 
\arrayrulecolor{lightgray}
\textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\\hline
1 &         CD+FS &    49  &  13 & \quart{7}{25}{17}{115} \\\hline  
2 &      CD &    59  &  18 & \quart{15}{34}{36}{115} \\
2 &          BIC &    60  &  12 & \quart{24}{24}{38}{115} \\\hline  
3 &      XTREE &    62  &  6 & \quart{36}{12}{42}{115}  \\\hline \end{tabular}}
\end{center}

In this example table, the rows are  sorted on the median values of each method. Note that all the methods
have   $R<0\%$ values; i.e. all these methods reduced the expected value of the performance score in that experiment
while XTREE achieved the greatest reduction (of 60\% from the original value).

The above eample table has a  left-hand-side  {\bf Rank} column, computed using the
Scott-Knott test described above. This column reports if
the values for each method are statistically different and are more than trivially different. 
In this example table, CD and BIC are ranked together while XTREE and CD+FS are ranked best and worst, respectively.

@@

\fig{jur} and \fig{conf1} show the effectiveness of our methods seen in 40 repeats with each data set.
In these experiments,   the dependent variables of Jureczko and Siegmund data sets are discrete and continuous in nature, respectively. Hence, while choosing the predictor, we used Random Forest (1) as a classifier for Jureczko data and (2) as a regressor for Siegmund data.
\bi
\item For Siegmund data, we performed a k-fold cross validation. Given the relatively smaller sample sizes in Apache and BDBJ, we chose $k=2$. This, we reasoned, would maintain a sufficiently large test data size in order for R to be measured reliably. However, in 2-fold cross validation, the division of data tends to affect the outcome significantly. To counter this, we randomized the order of the data, training on one half while identifying treatment plans on the remaining test data, repeating the process 40 times as mentioned above.
\item For the Jureczko data, we used the training and testing sets of \fig{j}. For these datasets,
 all the SMOTE-ing and Random Forest tunings (discussed in \tion{tesd})
occurred in the {\em train} phase of \fig{design}.
\ei
 \subsection{Experimental Results}

@@

Measured in terms of effectiveness,
some data sets were harder to optimize that others.  SQL (in \fig{conf1}) defied all 
our methods for reducing runtimes. Also, BDBJ (in \fig{conf1}) was hard to optimize
for all methods except XTREE.  However, in other data sets,
large reductions were observed:
\bi
\item
Down to 22\% of the original baseline in Ant of \fig{jur};
\item
Down to 6\% of the original baseline in BDBD  of \fig{conf1};
\ei
Overall, XTREE was most effective. It was always  the top-ranked method and (with 
the exception of SQL), had significant reductions in the median performance
values: median improvement of least 10\% lower (i.e. better) than the next ranked method.

@@

  In these plots, the {\em more} succinct a planning method, the {\em less} percent
  of the runs
  where 
  it   recommends changing a particular feature (i.e. the vertical bars in that plot
  are {\em lower}). For example,
 Method1 (CD) was the least succinct  since it  wanted to change all features 
 (observe the change frequencies as 
high as 100\% for all features). 
Method1's policy of ``change everything''  might be acceptable
if this approach lead 
to the most effective changes. However, looking at 
\fig{jur} and  \fig{conf1}, there is no evidence for this.

Overall,  XTREE's plans were usually succinct--  in all our data sets,
 XTREE changes 
 usually changes around a fifth of the features (see \fig{types}.

>>>>

 This paper proposes  the  XTREE planner for software projects. 
XTREE inputs tables of data of independent features plus a   weighted class
  indicating how much each row is  ``bad'' or  ``better''. 
  Plans are edits to  
rows such that the changed row is more likely to be ``better''.  XTREE
learns those plans by building a decision tree across the data, then reporting
the differences in the branches from some current branch to another desired branch.
Using data from 11 software projects,  XTREE can find  larger reductions than three
alternate methods.  Those reductions had 
median size of (56\%,28\%) and   largest size of (60\%,77\%) in
(defects,runtimes), respectively.
\end{abstract}
\section{Introduction}
Business users are demanding  tools that support    business-level interpretations of their data. At a panel on software analytics at ICSE`12, industrial practitioners lamented the state of the art in software analytics~\cite{menzies12a}. Panelists commented  ``prediction is all well and good, but what about decision making?''. Note that these panelists were more interested in the interpretations and follow-up that occurs after the mining, rather than just the mining itself. So:
\bi
\item
Instead of just accepting {\em predictions} on how many software defects to expect, business users might now demand a {\em plan} to reduce the likelihood of those defects.
\item
Instead of just accepting {\em predictions} on the runtime time of their software, business users might now demand a {\em plan} to reduce that runtime. 
\ei
In response to this business-level demands for planners,
we propose a novel {\em planning} method called XTREE for learning changes to a software system
such that its performance ``improves'', according to some measure. This paper uses XTREE
to reduce  the expected value of the
defects     in Jureczko et al.'s    JAVA systems~\cite{jureczko10};
and the runtimes   in   software    configured by  Siegmund et al.~\cite{sven12}.

@@

The rest of this paper   describes our data, our planners, and the experiment that
ranks XTREE against alternate approaches.  This is followed by notes on related work
and validity.
To enable reproducibility, all scripts and data used in this 
study are available online at \url{http://git.io/vG3DG}.

@@

\subsection{From Prediction to Planning}\label{sect:p2p}

@@

One way to propose changes to a project would be to   ask some smart experienced
person for their opinion on how to (e.g.) reduce defects and/or decrease runtimes. Sometimes
such advice 
is an effective strategy and sometimes it is not.
According to Passos et al.~\cite{passos11},  developers
may  assume that the lessons they learn from a few past
projects are general to 
all their future projects. They comment ``past experiences were taken into account without 
much consideration for their context''~\cite{passos11}.  
 J{\o}rgensen \& Gruschke~\cite{jorgensen09} offer  a similar warning. They report that the 
  supposed software engineering  ``Gurus'' rarely use lessons
  from past projects to improve their future reasoning and that such poor
  past advice can be detrimental to new projects.~\cite{jorgensen09}.
  Accordingly, we   propose a ``trust, but verify'' approach.
  After a software guru offers some sage wisdom,  it is wise to ask some other oracle 
  if there are any better options
  (just as a sanity check).
  The rest of this paper discusses some methods to build automatic oracles 
  to implement that   sanity check.

Another way  to find   changes to a project
might be to rely
on the peer review processes used by the 
SE research community. This approach would propose changes to software
projects that concur with internationally accept best practices. 
There are two
problems with that approach. Firstly, given the rapid pace of change in software
engineering, we may be asking questions for which there is no current widely
accepted concept of
``best practice''.

@@

\subsection{Trusting the Changes}\label{sect:trust}
   XTREE is evaluated by  comparing
predicted performance scores before and after a planner makes changes to the feature values of an example:
After making those
changes, we may have a new example that has never been seen before. Therefore, it must be asked
{\em ``can we trust the predictions made on such new examples?''}

@@

Accordingly, we need {\em trust-increasing} planners to generate changed examples {\em closer} to the
training examples.  To see how this works, 
 \fig{howxy} is from the {\em ivy} data
set (one of the Jureczko data sets used in this paper). It shows: (1)~the training examples in gray, (2)~the test examples in red, and (3)~the
changed  examples displaced after applying a plan (in green).
 Note that the  the   changed examples
cases  (shown in green)  fall closer to the training cases (shown in gray) than
the test cases (shown in red).

@@

 \section{Test Data}\label{sect:tesd}
To assess our planning methods, our data (see git.io/vGYxc) comes from Jureczko et al.'s object-oriented JAVA systems~\cite{jureczko10}  and
  software system   configuration data from by  Siegmund et al.~\cite{sven12}.

   The Jureczko data records the number of known defects for each class using a post-release bug tracking system. The classes are described in terms of
  nearly two dozen metrics such as number of children (noc), lines of code (loc), etc. For details on the Jureczko data, see  \fig{ck} and \fig{jd}. 
 All the planning methods of this paper reflect on the numeric value of the raw defect counts. The predictor considers defects to a Boolean class data set, where defects are \texttt{TRUE} if the numeric defect count is greater than zero and \texttt{FALSE} otherwise.

@@

  Our evaluation strategy (discussed below) divides this data into a training and a testing set.
  From the training set we apply a data miner (to learn a quality predictor) and
  various planning methods (to learn different plans). Next, we try applying
  each of those plans to the  test set and ask the quality predictor to assess the changed examples.
  Finally, we say that the  ``best'' planner is the one that most reduces the predicted values
  in the changed examples.

  As mentioned in the last section,  this approach depends on having effective predictors for assessing the results.
  For the Siegmund data, this criteria was   relatively easy to achieve.
  The data in those data sets have a continuous class (runtime of the compiled system)
  so the performance of a quality predictor can  be measured in terms of  difference between the predicted runtime $p$ of test case items and their actual runtimes $a$ using  $s= (1 - \frac{abs(a - p)}{a})\times100\%$ ({\em higher} values are {\em better}).
This paper  explores six Siegmund configuration data sets:  Berkeley DB (Java and C versions), Apache, SQLite, LLVM, and
  x264. 
  As a preliminary study, we split that data into equal sized train:test groups
  and trained a Random Forest
  Regressor (from the SciKit learn kit~\cite{Pedregosa2012})   on one half, then applied to the other. After repeating this process 40 times, we achieved nearly perfect accuracy of $s=\{99.9, 99.8, 99.4, 99.1, 96.1\}\%$;
  i.e.  we are confident that the predictors from the Siegmund data can assess
our plans. (Aside: if the reader doubts that such high scores are achievable, we note that these scores are consistent with those achieved by predictors built by Siegmund et al.~\cite{sven12}.)

@@

As a preliminary study, we split the Jureczko  data into train and test groups.
Random Forest (again, from the SciKit learn kit~\cite{Pedregosa2012}) were
built from the training data, then applied to the test data.
The ``untuned'' columns of \fig{j} shows those results. We called a data set ``usable'' if Random Forest was able to classify majority of the instances correctly. For this purpose, we set a threshold of $\mathit{pd}\ge 60 \wedge \mathit{pf} \le 30$\% to select suitable data sets. With this threshold, however, none of our data sets were be suitable for this study.

@@

In summary, while we cannot trust predictors from some of our Jureczko data sets,
we can plan ways to reduce defects in {\em jedit, ivy, ant, lucene} and {\em poi}.
Accordingly, when this study explores the Jureczko data, we will use these five data sets.

@@

This section describes XTREE (which we call Method4) and  three  
alternate methods for learning plans.
  XTREE  uses the decision tree learner of \fig{where}.D.  It is new to this paper.
  The other methods use the   top-down
	bi-clustering method described in \fig{where}.C  which recursively divides the
	data in two  using a dimension that captures the greatest variability in the data. 
	We proposed Method 1 and 2   in 2012~\cite{me12c} while Methods 3 comes from research conducted earlier this year~\cite{krishna15}. 
All  methods have the  properties proposed in \tion{prelim}: (1)~they are {\em local learners}
that find   plans   particular to each test case; and (2)~they  are {\em trust-increasing}; 
  i.e. they change  examples such that they move {\em closer} to the training data.

@@

Our  description of the methods adopts the following convention. All variables
  set via  our engineering judgement  with Greek letters; e.g. $\alpha,\beta,\gamma,\omega$.
  In this paper, we show our current settings to these variables produces useful
  results. Elsewhere\cite{krall14,fu:ase15}, we are exploring tuning methods to 
  find better settings but  we have nothing definitive yet to report
  on auto-tuning planners.

@@

{\em Assumption1: } Large data sets can be adequately represented by a few dozen (or so) 
centroids.

 {\em Details1:} Method1 clusters project data by reflecting   on the independent variables, then
  reports the delta between the cluster centroids. 
  After clustering   training data using the WHERE algorithm of  \fig{where}.C, Method1
replaces all clusters with a  centroid $C_i$ computed from the median value of each
continuous/discrete feature. After that, it
finds the closest centroid $C_j$ that has a better
performance score. For Jureczko data, ``better'' means fewer defective examples while for the Seigmund data,
``better'' means lower median runtimes for the examples in that cluster.
Method1 then caches the  delta between the independent features between $C_i$ and $C_j$. For continuous
features, this delta is $C_j - C_i$. For discrete values, this delta is the value of that feature
in $C_j$. 
Finally, for every test case, Method1 use the distance measure $d$ shown in \fig{where}.B to find
the nearest centroid $C_i$.  It then proposes a plan for improving that test case
that is the conjunction of all the deltas between $C_i$ and $C_j$.

@@

{\em Summary2:} Method2 works like Method1 but now the   plans  only
mention the $\beta=33\%$ most informative features. Hence, Method2's plans are simpler.

{\em Assumption2: }  When reasoning about centroids, we can
just use features that best distinguish centroids; i.e. whose values appear in just a few centroids.

@@

 \input{tex/algos}
\subsubsection{Method3= Best In Cluster (BIC)=   Method1 + Gradient}\label{sect:BIC}

{\em Summary3:} Method3 is like Method1, but it uses more knowledge about the training data.

{\em Assumption3: } There exists ``gradients'' between clusters which, if used, will better guide us to finding better plans.

{\em Details3:}
Method3 generates clusters from the training data just like Method1. Following this, it summarizes the clusters $C_X$ using ``Best-In-Cluster'' $B_X$, which is (1) the centroid of the cluster for the Jureczko data; (2) the cluster's member with the fastest runtime for the Seigmund data. Following this, Method3 connects  each cluster $C_i$ to a nearest neighbor $C_j$ by a {\em gradient}. Each gradient has a (bottom,top) end labelled  ($C_i,C_j$) containing the (worst,best) performance scores, respectively.

For each test instance, Method3 finds the nearest gradient, the runs up to the top (best) end $C_j$, then extracts $B_j$ (which is the best-in-cluster associated with  $C_j$). The returned plan is then computed as a fraction ($\omega=0.5$) of the distance between the test case and it's corresponding $B_j$.

@@

 \begin{figure}[!ht]
{\small 
\[
\begin{array}{r} 
\mathrm{project}\\
\mathrm{data}
\end{array} 
\left\{\begin{array}{l}\mathit{train}
        \left\{\begin{array}{l}
                \mathrm{learn\;a\;}\mathrm{predictor\;}\mathrm{(e.g.\;via \;Random\;Forest)}\\
                \mathrm{learn\;a\;}\mathrm{planner\;}\mathrm{(e.g.\;via \; XTREE)}
              \end{array}\right.
       \\
      ~\\
\mathit{test}  
    \left\{\begin{array}{l@{~}l}
           \mathit{before}& =\mathrm{Performance\; scores \hspace{2pt} in \hspace{2pt}}\mathit{test}\\
           \mathrm{\bf if\;}\mathit{before} & >  \mathit{0}\\
           \mathrm{\bf then} &
           \left\{
            \begin{array}{l}
                \mathit{test'} = \mathrm{planner}(\mathit{test})\\
                \mathit{after} =\mathrm{predictor}(\mathit{test'})\\ 
                \mathrm{{\bf return}\;} R=(1-\frac{\mathit{after}}{\mathit{before}})\times 100 \%
            \end{array}
          \right.
   \end{array}\right.
\end{array} \right. 
\]}
 \caption{Experimental design .}\label{fig:design}
 \end{figure}

@@

Else, we (1)~apply the { planner} to alter the {\em test} data;
then (2)~apply the { predictor} to the altered data $test'$;
then (3)~return data on the {\em before, after} predictions expressed as percent improvement, denoted by  $R=(1-\frac{\mathit{after}}{\mathit{before}})\times100\%$, with the following following properties:
\bi
\item If $R  = 0\%$, this means  ``no change from baseline''; 
\item If $R \gt 0\%$, this indicates ``some reduction to the baseline'';
\item If $R \lt 0\%$, this indicates ``optimization failure''.
\ei
 \subsubsection{Statistical Methods}
 Our methods use some stochastic algorithms; e.g. WHERE's selection of ``what example to explore first'' (see \fig{where}.C) and
  XTREE' occasional use of a random guess when deciding what part of a discretized range to include in the plan
  (see \fig{xtrees_bare}). Hence, we report the $R$ values seen in 40 repeated runs,
  with different random number seeds
(we use 40 since that is  more than the 30 samples  needed
to satisfy the central limit theorem).

  To rank our methods using the results from these 40
  repeats, we use the Scott-Knott test recommended by Mittas and  Angelis~\cite{mittas13}.

In accordance to that test, using the median values of each method, we
sort a list of  $l=40$ values of $R$ values found in  $ls=4$ different methods. 
Then, we
split $l$ into sub-lists $m,n$ in order to maximize the expected value of
 differences  in the observed performances
before and after divisions. E.g. for lists $l,m,n$ of size $ls,ms,ns$ where $l=m\cup n$:
 \[E(\Delta)=\frac{ms}{ls}abs(m.\mu - l.\mu)^2 + \frac{ns}{ls}abs(n.\mu - l.\mu)^2\]

We then apply a apply a statistical hypothesis test $H$ to check
if $m,n$ are significantly different  (in our case, the conjunction of A12 and bootstrapping). 
If so, Scott-Knott  recurses on the splits.
In other words, we divide the data if \textit{both} bootstrap sampling and effect size test agree that a division is statistically significant (with a confidence of 99\%) and not a small effect ($A12 \ge 0.6$).

@@

Our results are presented in the form of line diagrams like those shown on the right-hand-side of the following example table.
The black dot shows the median $R$ value and the horizontal likes stretch 
from the 25th percentile to the 75th percentile (a region called the inter-quartile
range, or IQR).

@@

{\small Example \begin{tabular}{{l@{~~~~}l@{~~~~}r@{~~~~}r@{~~}c@{}r}} 
\arrayrulecolor{lightgray}
\rowcolor{lightgray}\textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\\hline
1 &      XTREE &    62  &  6 & \quart{47.2}{4.8}{49.6}{115}  \\
\hline 2 &          BIC &    50  &  12 & \quart{34.2}{9.6}{40}{115} \\ 
\hline 3 &      CD &    44  &  18 & \quart{28}{14.4}{35.2}{115} \\
3 &         CD+FS &    43  &  13 & \quart{30}{10.4}{35}{115} \\  
\hline \end{tabular}}
\end{center}

In this example table, the rows are  sorted on the median values of each method. Note that all the methods
have   $R\gt0\%$ values; i.e. all these methods reduced the expected value of the performance score in that experiment
while XTREE achieved the greatest reduction (of 62\% from the original value).

The above example table has a  left-hand-side  {\bf Rank} column, computed using the
Scott-Knott test described above. In this example table, XTREE is ranked the best, while CD and CD+FS together are ranked the worst.

@@

\fig{jur} and \fig{conf1} show the effectiveness of our methods seen in 40 repeats with each data set.
In these experiments,   the dependent variables of Jureczko and Siegmund data sets are discrete and continuous in nature, respectively. Hence, while choosing the predictor, we used Random Forest (1) as a classifier for Jureczko data and (2) as a regressor for Siegmund data.

For Siegmund data, we performed a k-fold cross validation. Given the relatively small sample sizes in Apache and BDBJ, we chose $k=2$. This, we reasoned, would maintain a sufficiently large test data size in order for R to be measured reliably. However, in 2-fold cross validation, the division of data tends to affect the outcome significantly. To counter this, we randomized the order of the data, training on one half while identifying treatment plans on the other, repeating the process 40 times as mentioned above.

The Jureczko data, being temporal in nature, allows us
to implement a validation procedure that ensures only past data is ever used to
predict future values. Hence, in that data, we used the train/test sets shown in \fig{j}. 
(Aside: note  that all the SMOTE-ing and Random Forest tunings (discussed in \tion{tesd}) occurred in the {\em train} phase of \fig{design}).

\subsection{Experimental Results}

@@

Measured in terms of effectiveness,
some data sets were harder to optimize that others.  SQL (in \fig{conf1}) defied all 
our methods for reducing runtimes. While all methods performed equally well for Jedit, XTREE was the only method that could optimize BDBJ (in \fig{conf1}). In general, in most data sets,
large reductions were observed:
\bi
\item
An improvement of 60\%, as compared to the original baseline in Ant and Lucene of \fig{jur};
\item
An improvement of 77\% as compared to the original baseline in BDBC  of \fig{conf1};
\ei

Overall, XTREE was most the effective. It was always the top-ranked method with 
the exception of SQL and Jedit. Where it ranked better, it had significant improvements in the median performance values (1) In the Jureczko data sets, there was a median improvement of around 10\% larger than the next ranked method; (2) In the Seigmund data sets, improvements as large as 36\% greater than the next ranked method were observed.

@@

  In these plots, the {\em more} succinct a planning method, the {\em less} percent
  of the runs
  where 
  it   recommends changing a particular feature (i.e. the vertical bars in that plot
  are {\em lower}). For example,
  XTREE's plans were usually succinct--  in all  data sets,
 XTREE   changes around a fifth of the features (see \fig{types}).
On other hand,  Method1 (CD) was the least succinct  since it  wanted to change all features 
 (observe the change frequencies as 
high as 100\% for all features). 
Method1's policy of ``change everything''  might be acceptable
if this approach lead 
to the most effective changes. However, in
\fig{jur} and  \fig{conf1}, there is no evidence for this.
